Music Perception: Pilot Analysis

### 2021-01-22 analysis/analysis.Rmd
add stats
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: "Pilot Analysis (music perception study)"
author: "Atsuko Tominaga"
output: html_notebook
---

@@

- 10 participants (experimental error = 1)

```{r packages, include = FALSE}
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
```

@@

# Accuracy
## Individual data
```{r summary, echo = FALSE}
# remove irrelevant strings
data$midFile <- gsub("./practice/", "", data$midFile)
data$midFile <- gsub("./mid/", "", data$midFile)
data$midFile <- gsub(".mid", "", data$midFile)

@@

## Summary
### all
```{r acc, echo = FALSE}
# individual
ind_data <- data[expMode == "experiment" & subjectNumber != 9, .(N = .N, Sum = sum(correct), Percent = (sum(correct)/.N)*100), by = .(subjectNumber)]
ind_data <- ind_data[order(subjectNumber)]

# all
subset <- ind_data[, .(N = .N, Mean = mean(Sum), SD = sd(Sum), Mean_Percent = (mean(Sum)/96)*100, SD_Percent = (sd(Sum)/96)*100)]
subset
```
### articulation vs. dynamics

```{r plot, echo = FALSE}
ind_data_t <- data[expMode == "experiment" & subjectNumber != 9, .(N = .N, Sum = sum(correct), Percent = (sum(correct)/.N)*100), by = .(subjectNumber, technique)]
ind_data_t <- ind_data_t[order(subjectNumber, technique)]

subset_t <- ind_data_t[, .(N = .N, Mean = mean(Sum), SD = sd(Sum), Mean_Percent = (mean(Sum)/48)*100, SD_Percent = (sd(Sum)/48)*100), by = .(technique)]
subset_t

data_box <- data.table(subjectNumber = ind_data_t$subjectNumber, articulation = ind_data_t[technique == "articulation"]$Percent, dynamics = ind_data_t[technique == "dynamics"]$Percent)
box <- ggpaired(data_box, cond1 = "articulation", cond2 = "dynamics", line.color = "gray", line.size = 0.4, ylab = "Correct %")
box
```

## 1. teaching vs. performing (altogether)
```{r plot1, echo = FALSE}
# individual
ind_data1 <- data[expMode == "experiment" & subjectNumber != 9, .(N = .N, Sum = sum(correct), Percent = (sum(correct)/.N)*100), by = .(subjectNumber, condition)]
ind_data1 <- ind_data1[order(subjectNumber, condition)]

# all
subset1 <- ind_data1[, .(N = .N, Mean = mean(Sum), SD = sd(Sum), Mean_Percent = (mean(Sum)/48)*100, SD_Percent = (sd(Sum)/48)*100), by = .(condition)]
subset1

data_box1 <- data.table(subjectNumber = ind_data1$subjectNumber, teaching = ind_data1[condition == "teaching"]$Percent, performing = ind_data1[condition == "performing"]$Percent)
box1 <- ggpaired(data_box1, cond1 = "teaching", cond2 = "performing", line.color = "gray", line.size = 0.4, ylab = "Correct %")
box1
```

## 2. teaching vs. performing (each technique)
```{r plot2, echo = FALSE}
# individual
ind_data2 <- data[expMode == "experiment" & subjectNumber != 9, .(N = .N, Sum = sum(correct), Percent = (sum(correct)/.N)*100), by = .(subjectNumber, condition, technique)]
ind_data2 <- ind_data2[order(subjectNumber, condition, technique)]

# all
subset2 <- ind_data2[, .(N = .N, Mean = mean(Sum), SD = sd(Sum), Mean_Percent = (mean(Sum)/24)*100, SD_Percent = (sd(Sum)/24)*100), by = .(condition, technique)]
subset2

data_box2 <- data.table(subjectNumber = ind_data2$subjectNumber, teaching = ind_data2[condition == "teaching"]$Percent, performing = ind_data2[condition == "performing"]$Percent, technique = ind_data2$technique)
box2 <- ggpaired(data_box2, cond1 = "teaching", cond2 = "performing", facet.by = "technique", line.color = "gray", line.size = 0.4, ylab = "Correct %")
box2
```

## 3. teaching (each item)
```{r item1, echo = FALSE}
# individual
ind_data3 <- data[expMode == "experiment" & subjectNumber != 9 & condition == "teaching", .(N = .N, Sum = sum(correct), Percent = (sum(correct)/.N)*100), by = .(subjectNumber, condition, technique, item)]
ind_data3 <- ind_data3[order(subjectNumber, condition, technique, item)]

@@

## 3. performing (each item)
```{r item2, echo = FALSE}
# individual
ind_data4 <- data[expMode == "experiment" & subjectNumber != 9 & condition == "performing", .(N = .N, Sum = sum(correct), Percent = (sum(correct)/.N)*100), by = .(subjectNumber, condition, technique, item)]
ind_data4 <- ind_data4[order(subjectNumber, condition, technique, item)]

>>>>

---
title: "Music Perception: Pilot Analysis"
author: "Atsuko Tominaga"
output: html_notebook
---

@@

- 10 participants (experimental error = 1) >> 9 participants were included for analysis.

# Experiment

## Stimuli
- Stimuli were selected from the previous study ([the sound of teaching music/experiment-1](https://github.com/atsukotominaga/music-teaching/tree/main/experiment-1)).
- 48 stimuli were randomly chosen from each technique (either articulation or dynamics - therefore in total 96 stimuli). The half of them were selected from the teaching condition and the other half were from the performing condition.
  + 24 for teaching-articulation performances
  + 24 for performing-articulation performances
  + 24 for teaching-dynamics performances
  + 24 for performing-dynamics performances

## Design

- 2 blocks (articulation / dynamics): participants were asked to do the same task for two expressions separately. The order of the blocks was counterbalanced across participants.
- 48 trials for each block (each stimulus was presented only once) and the stimuli were randomised across participants.

## Procedure
1. Musicians (> 6yo experience) were asked to listen to a number of recordings and judge whether each performance was produced for teaching purposes or not.

Actual instruction:
```
Each performance was produced in order to either1) teach the musical expressive technique (as a teacher) 2) perform their best (as a performer).

You will be asked to judge whether each performer had the intention to teach or not by pressing the 'Yes' <Left> or 'No' <Right> key.
```

2. Participants pressed "YES" if they think the recording was produced for teaching (as a teacher) whereas they pressed "NO" if they thought the recording was produced for performing (as a performer).

3. Correct answers would be 1) participants pressed "YES" when they listened to teaching performances or 2) participants pressed "NO" when they listened to performing performances.

I hope these make sense...

```{r packages, include = FALSE}
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
if (!require("afex")) {install.packages("afex"); require("afex")}
```

@@

# Results
## Individual data
```{r summary, echo = FALSE}
# remove irrelevant strings
data$midFile <- gsub("./practice/", "", data$midFile)
data$midFile <- gsub("./mid/", "", data$midFile)
data$midFile <- gsub(".mid", "", data$midFile)

@@

## 1. Summary
### all
```{r acc, echo = FALSE}
# individual
ind_data <- data[expMode == "experiment" & subjectNumber != 9, .(N = .N, Sum = sum(correct), Percent = (sum(correct)/.N)*100), by = .(subjectNumber)]
ind_data <- ind_data[order(subjectNumber)]

# all
subset <- ind_data[, .(N = .N, Mean = mean(Sum), SD = sd(Sum), Mean_Percent = (mean(Sum)/96)*100, SD_Percent = (sd(Sum)/96)*100)]
subset
```

```{r plot, echo = FALSE}
ggboxplot(ind_data, x = "N", y = "Percent", add = "dotplot", xlab = FALSE) + ylab("Correct %") + theme(axis.text.x = element_blank())
```

### One-sample *t*-test
```{r t-test, echo = FALSE}
# normality check
shapiro.test(ind_data$Percent)

t.test(ind_data$Percent, mu = 50, alternative = "two.sided")
```

- Surprisingly, it showed a trend ( p = .06) but the 95% confidence interval overlaps with 50 %.

### articulation vs. dynamics

```{r plot2, echo = FALSE}
ind_data_ad <- data[expMode == "experiment" & subjectNumber != 9, .(N = .N, Sum = sum(correct), Percent = (sum(correct)/.N)*100), by = .(subjectNumber, technique)]
ind_data_ad <- ind_data_ad[order(subjectNumber, technique)]

subset_ad <- ind_data_ad[, .(N = .N, Mean = mean(Sum), SD = sd(Sum), Mean_Percent = (mean(Sum)/48)*100, SD_Percent = (sd(Sum)/48)*100), by = .(technique)]
subset_ad

data_box <- data.table(subjectNumber = ind_data_ad$subjectNumber, articulation = ind_data_ad[technique == "articulation"]$Percent, dynamics = ind_data_ad[technique == "dynamics"]$Percent)
box <- ggpaired(data_box, cond1 = "articulation", cond2 = "dynamics", line.color = "gray", line.size = 0.4, ylab = "Correct %")
box
```

## One-sample *t*-test
### articulation
```{r t-test-a, echo = FALSE}
# normality check
shapiro.test(ind_data_ad[technique == "articulation"]$Percent)

t.test(ind_data_ad[technique == "articulation"]$Percent, mu = 50, alternative = "two.sided")
```
- Not significantly different from chance (p > .2).

### dynamics
```{r t-test-d, echo = FALSE}
# normality check
shapiro.test(ind_data_ad[technique == "dynamics"]$Percent)

t.test(ind_data_ad[technique == "dynamics"]$Percent, mu = 50, alternative = "two.sided")
```

- The performance was significantly better than chance (p < .01).

## Paired *t*-test between articulation and dynamics
```{r p-t-test, echo = FALSE}
# normality check
shapiro.test(ind_data_ad[technique == "dynamics"]$Percent-ind_data_ad[technique == "articulation"]$Percent)

t.test(ind_data_ad[technique == "articulation"]$Percent, ind_data_ad[technique == "dynamics"]$Percent, paired = TRUE)
```

- There was a difference between articulation and dynamics in terms of accuracy.

## 2. teaching (each item)
### all
```{r item1, echo = FALSE}
# individual
ind_data3 <- data[expMode == "experiment" & subjectNumber != 9 & condition == "teaching", .(N = .N, Sum = sum(correct), Percent = (sum(correct)/.N)*100), by = .(subjectNumber, condition, technique, item)]
ind_data3 <- ind_data3[order(subjectNumber, condition, technique, item)]

@@

## 3. performing (each item)
### all
```{r item2, echo = FALSE}
# individual
ind_data4 <- data[expMode == "experiment" & subjectNumber != 9 & condition == "performing", .(N = .N, Sum = sum(correct), Percent = (sum(correct)/.N)*100), by = .(subjectNumber, condition, technique, item)]
ind_data4 <- ind_data4[order(subjectNumber, condition, technique, item)]
